Characterizing Stellar Streams by ML

### 2023-09-15 src/tex/ms.tex
[showyourwork] overleaf sync

\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{amsmath, amssymb}
\usepackage{gensymb}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{dsfont}
\usepackage[inline]{enumitem}
\usepackage{cleveref}
\usepackage{todonotes}

@@

    Stellar streams are the disrupted remnants of globular clusters and
    satellite galaxies, and their census has grown extensively in the Milky Way
    thanks to all-sky astrometric missions like {\it Gaia} and low-surface
    brightness photometric surveys (e.g., \citealt{2018ApJ...862..114S,
    2021ApJ...914..123I}).  The formation of streams can be traced back to a
    progenitor cluster, whose stars are gradually lost to the tidal field of the
    host galaxy and extend along a series of similar orbits producing tidal
    tails, which can sometimes span several tens of degrees across the sky. The
    orbital proximity of stars belonging to a stream makes them sensitive
    tracers of the mass distribution of galaxies, sourced by both the baryonic
    and dark matter components.

    The average properties of a stream (e.g., the mean position and velocity
    track of the tidal tails) can be used to reconstruct the global dark matter
    distribution of the stellar halo (e.g., \JN{cite}), while local variations
    in the stream track and surface density encode information about the
    stream's encounters with baryonic structures such as (e.g.) the stellar bar
    (\JN{cite}), and non-luminous structures like dark matter subhalos
    (\JN{cite}). Importantly, these galactic components span orders of magnitude
    in mass and spatial extent, highlighting the sensitivity of streams in
    constraining the structure of the galaxy.

@@

    We now introduce our method for characterizing stellar stream tracks and
    computing stream membership likelihoods from N-D phase-space measurements.
    % TODO: rephrase following sentence
    We develop a flexible Bayesian model framework, that allows arbitrary
    numbers of phase-space dimensions to be incorporated in the model and
    permits missing phase-space observations.  Below, we offer a concise
    overview of our approach, followed by a more in-depth technical explanation
    in subsequent sections.  We start, in \autoref{sub:method:framing_the_data},
    by transforming the astrometric data into a stream-oriented on-sky reference
    frame ($\phi_1, \phi_2, \parallax, \mu_{\phi_1}^*, \mu_{\phi_2}, v_r)$, for
    each stream. In this frame, the stream track lies along $\phi_1$, and the
    stream density may be described by distributions in the other parameters as
    injective functions of $\phi_1$.
    % TODO: ⬆ I think a few more sentences explaining why this is important,
    % feature variations, etc?  However, previously we discussed how stream will
    % vary along the track, with changes in the width and possible gaps.  Also
    % observational effects, e.g the distance gradient changes the on-sky width
    % (should this be mentioned earlier?)
    Likewise, the background of the Galactic field may be described with
    $\phi_1$-conditioned distributions.
    % TODO: ⬆ I think this should be expanded.
    Non-astrometric coordinates, e.g. photometry and metallicity, are also
    modelled along the stream and background as functions of $\phi_1$.

    The distributions for the stream and background are collected in a single
    mixture model that describes the entire field.  We take the parameters of
    the mixture model -- the mixing coefficients and component distribution
    parameters -- to be general (continuous) functions of the conditioning
    variable, $\phi_1$. This is done using feed-forward neural network that
    takes the conditioning variable as its input and output the parameter value.
    Mixture models of conditional probability distributions using neural
    networks are called Mixture Density Networks (MDNs). Since the model
    coefficients may evolve with $\phi_1$, MNDs are well suited to characterize
    streams (and the background) with all their variations in position, width,
    and linear density.

@@

    \subsection{Framing the Data} \label{sub:method:framing_the_data}

@@

        \begin{figure}
            \centering
            \includegraphics[scale=0.5]{static/model/DAG_StreamDensity.pdf}
            \caption{Probabilistic Graphical Model of the Mixture Density Network (MDN)}
            \todo[inline]{TODO:}
            \begin{enumerate}
                \item remake this example this with
                \url{https://github.com/daft-dev/daft}. See
                \url{https://en.wikipedia.org/wiki/Mixture_model} for examples.
            \end{enumerate}
            \label{fig:DAG}
        \end{figure}

@@

        The mixture coefficients $\qth{f}$ are normalized s.t.
        \begin{equation}
            \sum_{\mrm{q}} \qth{f}(\phi_1) = 1 \qquad \forall \phi_1
        \end{equation}
        In practice this is enforced by defining
        \begin{equation} \label{eq:mixture_weight_normalization}
            f_{b}(\phi_1) = 1 - \sum_{\rm{q}\in \mbs{I_Q}\backslash \{b\}} \qth{f}(\phi_1),
        \end{equation}
        the summation over all models except for some ``background" index $b \in
        \mbs{I_Q}$.  We discuss the mixture weights further in
        \autoref{sub:method:mixture_weight}.

@@

        It will prove convenient to distinguish the models for $q$ as the
        \textit{background} and \textit{stream} models, respectively. In this
        notation:
        \begin{align} \label{eq:stream_and_bkg_prob}
            \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})
            =& \phantom{+} \qquad f \phantom{+} \Scmp{P}(\nth{\mbs{w}}|\mbs{\theta}) \Spdf(\nth{\mbs{m}}|\mbs{\theta}) \\
            & + (1-f) \Bpdf(\nth{\mbs{w}}|\mbs{\theta}) \Bpdf(\nth{\mbs{m}}|\mbs{\theta}), \nonumber
        \end{align}
        where we have dropped the $(w), (m)$ superscript as the model types are
        evident from their inputs. We applied
        \autoref{eq:mixture_weight_normalization} to require only a single
        mixture coefficient.  Of course, if more components are included in the
        mixture then \autoref{eq:stream_and_bkg_prob} becomes the more general
        \autoref{eq:general_model}. However, this an implementation detail as
        any more complex model, e.g a model including a cold stream, extra-tidal
        features, the background stellar halo, and field globular clusters may
        be rearranged such that all stream-related components (cold or
        extra-tidal) and all other components (stellar halo and globular
        clusters) are grouped into multi-component \textit{stream} and
        \textit{background} models which themselves adhere to a
        \autoref{eq:general_model}-type relation.

@@

        Streams have a diversity of measurement coverage: several have full 6-D
        astrometrics as well as photometrics and chemical abundances available
        \citep[e.g.,][]{Koposov+2019, Antoja+2020, Li+2022}, while most have
        partial coverage, lacking some feature dimensions.
        % TODO: terminology ND-phase-space vs feature columns
        Most streams in \Gaia {\color{red} [cite]}, for instance, have proper
        motions but not radial velocities nor reliable parallax measurements.
        Followup surveys, like \dataarchive{S5} {\color{red} [cite]}, add many
        feature dimensions, however only for a select set of streams. Given the
        diversity of coverage it is important that the likelihood model --
        \autoref{eq:membership_prob} -- account for missing data.

@@

        The mixture weight $f_q(\phi_1)$ as used in \autoref{eq:general_model}
        is characterized as the fraction of stars belonging to the mixture
        component $q$ over an increment $[\phi_1, \phi_1 + \dif{\phi_1}]$. Thus,
        the mixture weight encodes information about the linear density of the
        stream, which has been shown to depend on substructure in the galaxy
        (CITE-x). The dependence of the mixture weight on the linear density is
        captured by the expression
        \begin{equation}\label{eq:fraction_param}
            f_q\left(w_{i,0}\right) = \frac{N_q\left(\phi_1\right)}{N_{\rm tot}\left(\phi_1\right)},
        \end{equation}
        where $N_q\left(\phi_1\right)$ is the number of stars in component $q$
        within an interval $[\phi_1, \phi_1 + \dif{\phi_1}]$, and $N_{\rm tot}$
        is the total number of stars in the same band (including those in
        component $q$). The linear  density for component $q$ is then entirely
        specified by the function $N_q(\phi_1)$.

        The mixture model specified in \S-X only models the weight parameter
        $f_q$, and not the components of the ratio in
        \autoref{eq:fraction_param}. However, the linear density $N_q(\phi_1)$
        can be readily obtained from $f_q$, provided that $N_{\rm
        tot}\left(\phi_1\right)$ can be calculated. Importantly, this function
        does not depend on the parameters of the on-stream or off-stream model.
        Instead, $N_{\rm tot}(\phi_1)$ is simply the number of stars in a small
        $\phi_1$ increment. We therefore model $N_{\rm tot}\left(\phi_1\right)$
        as a post-processing step using a single normalzing flow in $\phi_1$.

        In particular, for a given field the distribution of $\phi_1$
        coordinates is specified by the set $\{\phi_1\}_f$. A normalizng flow is
        trained to this dataset, and will satisfy $\int d\phi_1
        P_{\phi_f}(\phi_1) = 1$ by construction. In order to re-weight the
        normalizing flow to obtain $N_{\rm tot}(\phi_1)$, we simply rescale
        $P_{\phi}(\phi_1)$ by $N_{\rm tot,0}$, which is the total number of
        stars in the field considered. We can then obtain the linear stream
        density with
        \begin{equation}
            N_q\left(\phi_1\right) = f_q N_{\rm tot} \left(\phi_1\right)  = f_q N_{\rm tot,0} P_{\phi_1}\left(\phi_1\right) .
        \end{equation}

@@

            The astrometric track is modeled with a neural network, encompassing
            each phase space dimension, parameterized as a function of $\phi_1$.
            The neural network output for a single stream component is the mean
            location of the stream track in each dimension,
            $\astroM{\mbs{\mu}}(\phi_1)$, and its associated covariance matrix
            $\astroM{\mbs{\Sigma}}(\phi_1)$. The probability for the data is
            then
            \begin{equation}
                \cmp{S,w}{\pdf}(\nth{\mbs{w}} | \mbs{\theta}) \! = \mcal{N}_F \big(\nth{\mbs{w}} | \astroM{\mbs{\mu}}\!(\phi_1), {\cmp{w}{\nth{\mbs{\Sigma}}}}^2 \! + \!  \astroM{\mbs{\Sigma}}^2\!\left(\phi_1\right) \!\big),
            \end{equation}
            where $\astroM{\mbs{\theta}}$ is the set of neural network
            parameters that define $\astroM{\mbs{\mu}}$ and
            $\astroM{\mbs{\Sigma}}$; and $\cmp{w}{\nth{\mbs{\Sigma}}}$ the data
            uncertainties.  In the simplest case, we take
            $\cmp{w}{\nth{\mbs{\Sigma}}}, \astroM{\mbs{\Sigma}}$ to be diagonal,
            consisting of both intrinsic dispersion for each dimension of
            $\mbs{w}$ and the data uncertainties added in quadrature. The data
            uncertainties $\cmp{w}{\nth{\mbs{\Sigma}}}$ are drawn from
            $\mbs{\delta}\mbf{W}$, the dataset of astrometric uncertainties (see
            \autoref{app:notation}).

@@

            Normalizing flows provide a flexible description of a probability
            density, by transforming samples generated from a simple base
            distribution (i.e., a Gaussian) to the more complicated target
            distribution. The transformations are typically non-linear outputs
            of a neural neural network, whose parameters are optimized until the
            target density accurately characterizes the data. The neural network
            transformation is differentiable, so that jacobian factors can be
            combined to ensure that the the target distribution remains a valid
            probability density (i.e., integrates to one) \TODO{[cite]}.

@@

        The custom model framework combines for each model a neural network with
        Bayesian probability functions. The latter are set by the model type,
        e.g. the Gaussian model has a Gaussian likelihood function, but the
        former, the neural network, is set and determined by the user. So long
        as the network has the correct number of inputs (e.g. one for $\phi_1$)
        and outputs (one per parameter), any feed-forward network may be used.
        Our default choice, and for which we provide a helper function, is an
        MLP with a sequence of \texttt{Linear} \!\!\textrightarrow \texttt{Tanh}
        \!\!\textrightarrow \texttt{Dropout} (hereafter \texttt{LinTanh})
        blocks. This is illustrated in \autoref{table:mlp_structure}.
        % FIXME! reference not being picked up correctly.
        Dropout layers are used to prevent network co-adaptation and
        over-fitting. Dropout works by randomly zeroing out nodes in the
        network, decreasing the importance of any single node, encouraging
        instead emergent patterns across the network.  One model for which the
        lin-tanh-dropout network cannot be used is the normalizing flow used in
        the photometric background model
        (\autoref{sub:method:photometric_model:off_stream}). Here we use the
        \package{nflows} package, with a conditional diagonal normal base
        distribution and composite transform network of a reverse permutation
        and masked affine auto-regressive transform layers. As noted in our data
        availability statement, all the codes used in this paper are bundled by
        \href{https://github.com/showyourwork/showyourwork}{\package{showyourwork}}
        in a public repository at \url{\GitHubURL{}}. We will describe for each
        stream model specifics relevant to that stream but refer readers to the
        publicly-available code itself for code implementation details.
        % TODO: Maybe more details in the Appendix?

@@

\section{Results: GD-1} \label{sec:results_gd1}

    Having successful applied the density modeling framework to a mock stream,
    we now apply the method to the stellar stream \stream{GD-1}.
    % TODO: some reiteration on the importance of GD-1
    We discuss the data selection in \autoref{sub:results_gd1:data}, model
    specifics in \autoref{sub:results_gd1:model}, and the results in
    \autoref{sub:results_gd1:results}.

@@

        We utilize data from \Gaia{} DR3 for the astrometric coordinates of each
        star.  To obtain more accurate color information, we cross-match the
        \Gaia{} field on \PanStarrs{} \citep{PanSTARRS1} using \Gaia's provided
        \texttt{Best-neighbors} catalog.  To limit the field to a region known
        to contain \stream{GD-1} we adopt broad cuts on the cross-matched data.
        We perform these rough cuts by requiring the data lie within the
        phase-space cube:
        \begin{itemize}
            \item $\phi_1 \in [-100, 40] \ \rm{\degree}$,
            \item $\phi_2 \in [-9, 5] \rm{\degree}$,
            \item $\parallax \in [-10, 1] \ \rm{mas}$,
            \item $\rm{bp-rp} \in [-1, 3] \ \rm{mag}$
            \item $g, r \in [0, 50] \ \rm{mag}$, and not \texttt{null}.
        \end{itemize}
        where we use \package{gala}'s \citep{gala, galav1.3}
        \code{GD1Koposov10} frame \citep{Koposov+2010} to define $\phi_1$,
        $\phi_2$, and associated proper motions.  For downloading purposes we
        split the query into 10 sub-regions in even $\phi_1$ increments. A plot
        of the $\phi_1, \phi_2$ regions is included in the
        \href{\GitHubURL{}}{\package{ShowYourWork}} repository.  Stream-specific
        coordinates ($\phi_1, \phi_2$) are not included in \Gaia.  One could
        embed the coordinate transform in the ADQL \citep{ADQL2.0}, or define
        great-circle arcs in ICRS \citep{ICRS1997}. Since only rough cuts are
        required and since the sub-region boxes are sufficiently small for a
        reasonable flat-sky approximation  we simply translate the ($\phi_1,
        \phi_2$) boxes to ICRS.  We correct the \PanStarrs{} photometry for dust
        extinction using the \textit{Bayestar19} \citep{Green+2019} models in
        the \package{dustmaps} package \citep{Green2018}.  \textit{Bayestar19}
        is a three-dimensional dust map of the Milky Way for which we use
        \Gaia's parallax distance information. For stars with missing parallax measurements we assume a distance of $8.5 \ \unit[kpc]$.  We include the $1-\sigma$ errors
        as a source of uncertainty in $\mbs{\Sigma}_n$, the per-star photometric
        uncertainty.
        We also correct the parallax for the zero-point using the \code{gaiadr3-zeropoint} package, which implements the findings from \cite{Lindegren+2021}.
        The result is a field populated by
        \variable{output/gd1/ndata_variable.txt} stars, from which we would like
        to identify stream members.

@@

\section{Palomar 5} \label{sec:results_pal5}

@@

>>>>

\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{amsmath, amssymb}
\usepackage{gensymb}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{dsfont}
\usepackage[inline]{enumitem}
\usepackage{cleveref}
\usepackage{transparent}
\usepackage{todonotes}

@@

    Stellar streams are the disrupted remnants of globular clusters and
    satellite galaxies, and their census has grown extensively in the Milky Way
    thanks to all-sky astrometric missions like {\it Gaia} and low-surface
    brightness photometric surveys (e.g., \citealt{2018ApJ...862..114S,
    2021ApJ...914..123I} \JN{add more}).  The formation of streams can be traced back to a
    progenitor cluster, whose stars are gradually lost to the tidal field of the
    host galaxy and extend along a series of similar orbits producing tidal
    tails, which can sometimes span several tens of degrees across the sky. The
    orbital proximity of stars belonging to a stream makes them sensitive
    tracers of the mass distribution of galaxies, sourced by both the baryonic
    and dark matter components.

    The average properties of a stream (e.g., the mean position and velocity
    track of the tidal tails) can be used to reconstruct the global dark matter
    distribution of the stellar halo (e.g., \citealt{1999ApJ...512L.109J, 2014ApJ...795...94B, 2014ApJ...795...95B,2022ApJ...940...22N,2023MNRAS.521.4936K}), while local variations
    in the stream track and surface density encode information about the
    stream's encounters with baryonic structures such as the stellar bar
   (e.g., \citealt{2017MNRAS.470...60E, 2017NatAs...1..633P}) , and non-luminous structures like dark matter subhalos (e.g., \citealt{2017MNRAS.466..628B,2019ApJ...880...38B}). Importantly, these galactic components span orders of magnitude
    in mass and spatial extent, highlighting the sensitivity of streams in
    constraining the structure of the galaxy.

@@

    We now introduce our method for characterizing stellar stream tracks and
    computing membership probabilities from kinematic and photometric measurements of stars. 
    We develop a flexible Bayesian modeling framework that incorporates all of the available kinematic information for a field of stars, including those with only partial phase-space measurements. Below, we offer a concise
    overview of our approach, followed by a more in-depth technical explanation
    in subsequent sections.  We start, in \autoref{sub:method:framing_the_data},
    by transforming the astrometric data into a stream-oriented on-sky reference
    frame ($\phi_1, \phi_2, \parallax, \mu_{\phi_1}^*, \mu_{\phi_2}, v_r)$, for
    each stream. In this frame, the stream track lies along $\phi_1$, and the
    stream density may be described by distributions in the other parameters as
    injective functions of $\phi_1$.
    % TODO: ⬆ I think a few more sentences explaining why this is important,
    % feature variations, etc?  However, previously we discussed how stream will
    % vary along the track, with changes in the width and possible gaps.  Also
    % observational effects, e.g the distance gradient changes the on-sky width
    % (should this be mentioned earlier?)
    Likewise, the background of the Galactic field may be described with
    $\phi_1$-conditioned distributions.
    % TODO: ⬆ I think this should be expanded.
    Non-astrometric coordinates, e.g. photometry and metallicity \JN{do we ever actually do [Fe/H]$(\phi_1)$? If not, can remove metallicity}, are also
    modelled along the stream and background as functions of $\phi_1$.

    The distributions for the stream and background are collected in a single
    mixture model that describes the entire field.  We take the parameters of
    the mixture model -- the mixture weights and component distribution
    parameters -- to be general (continuous) functions of the conditioning
    variable, $\phi_1$. This is done using a feed-forward neural network that
    takes the conditioning variable as its input and outputs the parameter value(s) that characterize the distribution of stars at a $\phi_1$ ``slice".
    Mixture models of conditional probability distributions using neural
    networks are called Mixture Density Networks (MDNs) \JN{cite CM Bishop}. Since the model
    coefficients may evolve with $\phi_1$, MNDs are well suited to characterize
    streams (and the background) with all their variations in position, width,
    and linear density.

@@

    \subsection{Framing the Data} \label{sub:method:framing_the_data}
        \JN{I like this section since we're really thinking deeply about ``what does it mean to construct a stream frame", which is important for this paper! Though I think the point can be made with less words and notation. I.e., I think we just want to establish that (a) streams can be parameterized as a function of some affine parameter, and (b) in a heliocentric frame even a stream on a circular orbit will not appear flat in on-sky angular coordinates. }

        \JN{My modification of the paragraph below:  As a stream progenitor orbits the Galactic center of mass, its path can be approximated locally as an elliptical segment. The associated
        arms of the stream generally align closely with its orbital plane. As viewed from a Galactocentirc reference frame, the stream assumes the form of a great circle enabling a 1:1 mapping from some phase parameter to a unique position along the stream. }

        \JN{
        While it would be ideal to characterize streams in a Galactocentirc frame, transformation to this frame requires knowledge of the solar motion around the galactic center, and accurate distances to a field stars (among other uncertain quantities). In a heliocentric frame, projection effects can obscure the shape of a stream leading to a non-trivial stream track.}

        \JN{
        In a heliocentric frame, projection effects can obscure this simple description. Thankfully... (as before)}

@@

        % \begin{figure}
        %     \centering
        %     \includegraphics[scale=0.5]{static/model/DAG_StreamDensity.pdf}
        %     \caption{Probabilistic Graphical Model of the Mixture Density Network (MDN)}
        %     \todo[inline]{TODO:}
        %     \begin{enumerate}
        %         \item remake this example this with
        %         \url{https://github.com/daft-dev/daft}. See
        %         \url{https://en.wikipedia.org/wiki/Mixture_model} for examples.
        %     \end{enumerate}
        %     \label{fig:DAG}
        % \end{figure}
        \begin{figure}
            \script{gd1/plot/pgm_stream.py}
            \centering
            \includegraphics[scale=0.5]{figures/gd1/pgm_stream.pdf}
            \caption{Probabilistic Graphical Model of the Mixture Density Network (MDN)}
            \label{fig:PGM_stream}
        \end{figure}

        \begin{figure}
            \script{gd1/plot/pgm_background.py}
            \centering
            \includegraphics[scale=0.5]{figures/gd1/pgm_background.pdf}
            \caption{Probabilistic Graphical Model of the Mixture Density Network (MDN) \JN{I think these could go side by side and take up the top of this page nicely (two column figure)}}
            \label{fig:PGM_background}
        \end{figure}

@@

        The mixture coefficients $\qth{f}$ are normalized such that
        \begin{equation}
        \label{eq:mixture_weight_sum}
            \sum_{\mrm{q}} \qth{f}(\phi_1) = 1 \qquad \forall \phi_1
        \end{equation}
        In practice this is enforced by defining a background
        \begin{equation} \label{eq:mixture_weight_normalization}
            f_{b}(\phi_1) = 1 - \sum_{\rm{q}\in \mbs{I_Q}\backslash \{b\}} \qth{f}(\phi_1),
        \end{equation}
        the summation over all models except for some ``background" index $b \in
        \mbs{I_Q}$.  We discuss the mixture weights further in
        \autoref{sub:method:mixture_weight}. \JN{I think it's enough just to have Eq.~\ref{eq:mixture_weight_sum}.}

@@

        It will prove convenient to distinguish the models for $q$ as the
        \textit{background} and \textit{stream} models, respectively. In this
        notation:
        \begin{align} \label{eq:stream_and_bkg_prob}
            \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})
            =& \phantom{+} \qquad f \phantom{+} \Scmp{P}(\nth{\mbs{w}}|\mbs{\theta}) \Spdf(\nth{\mbs{m}}|\mbs{\theta}) \\
            & + (1-f) \Bpdf(\nth{\mbs{w}}|\mbs{\theta}) \Bpdf(\nth{\mbs{m}}|\mbs{\theta}), \nonumber
        \end{align}
        where we have dropped the $(w), (m)$ superscript as the model types are
        evident from their inputs. We applied
        \autoref{eq:mixture_weight_normalization} to require only a single
        mixture coefficient.  \JN{I think we should simplify the sentences below, it interrupts the flow. For this work, can we just say: ``Note that the stream and background PDFs can consist of further components. This will become important for our characterization of the stream \stream{GD-1}, which appears spatially bifurcated.}Of course, if more components are included in the
        mixture then \autoref{eq:stream_and_bkg_prob} becomes the more general
        \autoref{eq:general_model}. However, this an implementation detail as
        any more complex model, e.g a model including a cold stream, extra-tidal
        features, the background stellar halo, and field globular clusters may
        be rearranged such that all stream-related components (cold or
        extra-tidal) and all other components (stellar halo and globular
        clusters) are grouped into multi-component \textit{stream} and
        \textit{background} models which themselves adhere to a
        \autoref{eq:general_model}-type relation.

@@

        Streams have a diversity of measurement coverage: several have full 6-D
        astrometrics as well as photometrics and chemical abundances available
        \citep[e.g.,][]{Koposov+2019, Antoja+2020, Li+2022}, while most have
        partial coverage, lacking some feature dimensions.
        % TODO: terminology ND-phase-space vs feature columns
        Most streams in \Gaia {\color{red} [cite]}\JN{Should be italics right?}, for instance, have proper
        motions but not radial velocities nor reliable parallax measurements.
        Followup surveys, like \dataarchive{S5} {\color{red} [cite]}, add many
        feature dimensions, however only for a select set of streams. Given the
        diversity of coverage it is important that the likelihood model --
        \autoref{eq:membership_prob} -- account for missing data.

@@

        The mixture weight $f_q(\phi_1)$ as used in \autoref{eq:general_model}
        is characterized as the fraction of stars belonging to the mixture
        component $q$ over an increment $[\phi_1, \phi_1 + \dif{\phi_1}]$. Thus,
        the mixture weight encodes information about the linear density of the
        stream, which has been shown to depend on substructure in the galaxy
        (CITE-x). The dependence of the mixture weight on the linear density is
        captured by the expression
        \begin{equation}\label{eq:fraction_param}
            f_q\left(\phi_1\right) = \frac{N_q\left(\phi_1\right)}{N_{\rm tot}\left(\phi_1\right)},
        \end{equation}
        where $N_q\left(\phi_1\right)$ is the number of stars in component $q$
        within an interval $[\phi_1, \phi_1 + \dif{\phi_1}]$, and $N_{\rm tot}$
        is the total number of stars in the same band (including those in
        component $q$). The linear  density for component $q$ is then entirely
        specified by the function $N_q(\phi_1)$.

        The mixture model specified in \S-X only models the weight parameter
        $f_q$, and not the components of the ratio in
        \autoref{eq:fraction_param}. However, the linear density $N_q(\phi_1)$
        can be readily obtained from $f_q$, provided that $N_{\rm
        tot}\left(\phi_1\right)$ can be calculated. Importantly, this function
        does not depend on the parameters of the on-stream or off-stream model.
        Instead, $N_{\rm tot}(\phi_1)$ is simply the number of stars in a small
        $\phi_1$ increment. We therefore model $N_{\rm tot}\left(\phi_1\right)$
        as a post-processing step using a single normalizing flow in $\phi_1$.

        In particular, for a given field the distribution of $\phi_1$
        coordinates is specified by the set $\{\phi_1\}_f$. A normalizing flow is
        trained to this dataset, and will satisfy $\int d\phi_1
        P_{\phi_f}(\phi_1) = 1$ by construction. In order to re-weight the
        normalizing flow to obtain $N_{\rm tot}(\phi_1)$, we simply rescale
        $P_{\phi}(\phi_1)$ by $N_{\rm tot,0}$, which is the total number of
        stars in the field considered. We can then obtain the linear stream
        density with
        \begin{equation}
            N_q\left(\phi_1\right) = f_q N_{\rm tot} \left(\phi_1\right)  = f_q N_{\rm tot,0} P_{\phi_1}\left(\phi_1\right) .
        \end{equation}

@@

            The astrometric track is modeled with a neural network, encompassing
            each phase space dimension, parameterized as a function of $\phi_1$.
            The neural network output for a single stream component is the mean
            location of the stream track in each dimension,
            $\astroM{\mbs{\mu}}(\phi_1)$, and its associated covariance matrix
            $\astroM{\mbs{\Sigma}}(\phi_1)$. The probability for the data is
            then
            \begin{equation}
                \cmp{S,w}{\pdf}(\nth{\mbs{w}} | \mbs{\theta}) \! = \mcal{N}_F {\transparent{0.5}\big(}\nth{\mbs{w}} | \astroM{\mbs{\mu}}\!(\phi_1), {\cmp{w}{\nth{\mbs{\Sigma}}}}^2 \! + \!  \astroM{\mbs{\Sigma}}^2\!\left(\phi_1\right) \!\big),
            \end{equation}
            where $\astroM{\mbs{\theta}}$ is the set of neural network
            parameters that define $\astroM{\mbs{\mu}}$ and
            $\astroM{\mbs{\Sigma}}$; and $\cmp{w}{\nth{\mbs{\Sigma}}}$ the data
            uncertainties.  In the simplest case, we take
            $\cmp{w}{\nth{\mbs{\Sigma}}}, \astroM{\mbs{\Sigma}}$ to be diagonal,
            consisting of both intrinsic dispersion for each dimension of
            $\mbs{w}$ and the data uncertainties added in quadrature. The data
            uncertainties $\cmp{w}{\nth{\mbs{\Sigma}}}$ are drawn from
            $\mbs{\delta}\mbf{W}$, the dataset of astrometric uncertainties (see
            \autoref{app:notation}).

@@

            Normalizing flows provide a flexible description of a probability
            density, by transforming samples generated from a simple base
            distribution (i.e., a Gaussian) to the more complicated target
            distribution. The transformations are typically non-linear outputs
            of a neural neural network, whose parameters are optimized until the
            target density accurately characterizes the data. The neural network
            transformation is differentiable, so that jacobian factors can be
            combined to ensure that the the target distribution remains a valid
            probability density (i.e., integrates to one) \TODO{[cite]}.
            For the neural networks we regularize the data (minus the mean, dividied by the standard deviation per feature column); thus we correct the overall normalization of the flow by the product of the jacobian of the regularization operation.

@@

        The custom model framework combines for each model a neural network with
        Bayesian probability functions. The latter are set by the model type,
        e.g. the Gaussian model has a Gaussian likelihood function, but the
        former, the neural network, is set and determined by the user. So long
        as the network has the correct number of inputs (e.g. one for $\phi_1$)
        and outputs (one per parameter), any feed-forward network may be used.
        Our default choice, and for which we provide a helper function, is an
        MLP with a sequence of \texttt{Linear} \!\!\textrightarrow \texttt{Tanh}
        \!\!\textrightarrow \texttt{Dropout} (hereafter \texttt{LinTanh})
        blocks. This is illustrated in \autoref{table:mlp_structure}.
        % FIXME! reference not being picked up correctly.
        Dropout layers are used to prevent network co-adaptation and
        over-fitting. Dropout works by randomly zeroing out nodes in the
        network, decreasing the importance of any single node, encouraging
        instead emergent patterns across the network.  One model for which the
        lin-tanh-dropout network cannot be used is the normalizing flow used in
        the photometric background model
        (\autoref{sub:method:photometric_model:off_stream}). Here we use the
        \package{zuko} package, with a conditional diagonal normal base
        distribution and composite transform network of a reverse permutation
        and masked affine auto-regressive transform layers. As noted in our data
        availability statement, all the codes used in this paper are bundled by
        \href{https://github.com/showyourwork/showyourwork}{\package{showyourwork}}
        in a public repository at \url{\GitHubURL{}}. We will describe for each
        stream model specifics relevant to that stream but refer readers to the
        publicly-available code itself for code implementation details.
        % TODO: Maybe more details in the Appendix?

@@

\section{Results: GD-1} \label{sec:results_gd1}
    We now apply our method for stream characterization to the stellar stream \stream{GD-1} , initially discovered by \citet{2006ApJ...643L..17G}. Previous work has identified several density variations along \stream{GD-1}, including a bifurcation that is well-modeled by an encounter with a dark matter subhalo. Characterizing the detailed density fluctuations of this stream is therefore important for constraining a population of perturbers.
    % TODO: some reiteration on the importance of GD-1
    We discuss the data selection in \autoref{sub:results_gd1:data}, model
    specifics in \autoref{sub:results_gd1:model}, and the results in
    \autoref{sub:results_gd1:results}.

@@

        We utilize data from \Gaia{} DR3 for the astrometric coordinates of each
        star.  To obtain more accurate color information, we cross-match the
        \Gaia{} field on \PanStarrs{} \citep{PanSTARRS1} using \Gaia's provided
        \texttt{Best-neighbors} catalog.  To limit the field to a region known
        to contain \stream{GD-1} we adopt broad cuts on the cross-matched data.
        We perform these rough cuts by requiring the data lie within the
        phase-space cube:
        \begin{itemize}
            \item $\phi_1 \in [-100, 40] \ \rm{\degree}$,
            \item $\phi_2 \in [-9, 5] \rm{\degree}$,
            \item $\parallax \in [-10, 1] \ \rm{mas}$,
            \item $\rm{bp-rp} \in [-1, 3] \ \rm{mag}$
            \item $g, r \in [0, 50] \ \rm{mag}$, and not \texttt{null}.
        \end{itemize}
        where we use \package{gala}'s \citep{gala, galav1.3}
        \code{GD1Koposov10} frame \citep{Koposov+2010} to define $\phi_1$,
        $\phi_2$, and associated proper motions.  For downloading purposes we
        split the query into 32 sub-regions in even $\phi_1$ increments.
        % TODO: make this a \variable
        A plot of the $\phi_1, \phi_2$ regions is included in the
        \href{\GitHubURL{}}{\package{ShowYourWork}} repository.  Stream-specific
        coordinates ($\phi_1, \phi_2$) are not included in \Gaia.  One could
        embed the coordinate transform in the ADQL \citep{ADQL2.0}, or define
        great-circle arcs in ICRS \citep{ICRS1997}. Since only rough cuts are
        required and since the sub-region boxes are sufficiently small for a
        reasonable flat-sky approximation  we simply translate the ($\phi_1,
        \phi_2$) boxes to ICRS.  We correct the \PanStarrs{} photometry for dust
        extinction using the \textit{Bayestar19} \citep{Green+2019} models in
        the \package{dustmaps} package \citep{Green2018}.  \textit{Bayestar19}
        is a three-dimensional dust map of the Milky Way for which we use
        \Gaia's parallax distance information. For stars with missing parallax measurements we assume a distance of $8.5 \ \unit[kpc]$.  We include the $1-\sigma$ errors
        as a source of uncertainty in $\mbs{\Sigma}_n$, the per-star photometric
        uncertainty.
        We also correct the parallax for the zero-point using the \code{gaiadr3-zeropoint} package, which implements the findings from \cite{Lindegren+2021}.
        The result is a field populated by
        \variable{output/gd1/ndata_variable.txt} stars, from which we would like
        to identify stream members.

@@

\clearpage
\section{Palomar 5} \label{sec:results_pal5}

@@

        \begin{figure}
            \script{pal5/plot/photometric_background_selection.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/pal5/photometric_background_selection.pdf}
            \caption{Caption}
            \label{fig:pal5-photometric_background_selection}
        \end{figure}

        % TODO: make this dynamically generated from the actual data.
        \begin{table}
            \centering
            \setlength{\tabcolsep}{0pt}
            \newcommand\capitem{\\$\phantom{+}\ast$\ }
            \caption{%
                Stream Track Regions Prior: %
                This table includes all the region priors used to guide the model towards the known stream track. The model will converge to the
                region $_{\rm minimum}^{\rm maximum}$.
                See \autoref{sub:methods:priors:track_region_prior} for details.
            }
            \label{tab:pal5_track_prior}
            {\rowcolors{2}{gray!10}{white!10}
            \begin{tabular}{@{}*{4}{r<{\hspace{7pt}}}l<{\hspace{7pt}}@{}}
                \toprule
                $\phi_1 \,\unit{\degree}$ & $\phi_2 \,\unit{\degree}$ & $\mu_{\phi_1} \,[\frac{\rm{mas}}{\rm{yr}}]$ & $\mu \,\unit{mag}$ & $(\simeq \parallax \,\unit{mas})$ \\
                \midrule
                0 & & & & \\ 
                \bottomrule\bottomrule
            \end{tabular}
            }
        \end{table}
